\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-07 paper.qmd
Updated PBMC3k section

As shown in @fig-NLDR-variety, layout c, d, and e that show small separations between clusters are universally optimal. But layout d with little separations perform well at tiny binwidth and poorly as binwidth increases. Also, layout e performs well at large binwidths. The layout c is chosen as the best, since it used the UMAP with a different (hyper-) parameter settings.

The visualization of the model generated for the selected layout in $9\text{-}D$ shows edges between the clusters (@fig-mnist-tri-proj (b2)). This provides evidence for the small separations between clusters. Furthermore, the data does not uniformally distributed along the clusters and therefore shows densed points (@fig-mnist-tri-proj (b2). Also, the clusters are non-linear shaped (@fig-mnist-tri-proj (b3).

>>>>

As a result, it is necessary to find an appropriate layout for the dataset. MSE for different binwidths ($a_1$) using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with various (hyper-)parameter settings were computed (@fig-pbmc-mse). Layouts c, d, and e, which show small separations between clusters, are universally optimal. However, layout d performs well with smaller binwidths and poorly with larger binwidths. On the other hand, layout e performs well with larger binwidths. Layout c was selected for further analysis due to its better (hyper-)parameter selection for the same method of the published plot.

The visualization of the selected layout in the $9\text{-}D$ shows edges between the clusters (@fig-mnist-tri-proj (b2)). This supports the presence of small separations between clusters. Additionally, the data points are not uniformly distributed across the clusters, resulting in dense areas (@fig-mnist-tri-proj (b2)). Furthermore, the clusters shows non-linear shapes (@fig-mnist-tri-proj (b3)).

<!-- As shown in @fig-NLDR-variety, layout c, d, and e that show small separations between clusters are universally optimal. But layout d with little separations perform well at tiny binwidth and poorly as binwidth increases. Also, layout e performs well at large binwidths. The layout c is chosen as the best, since it used the UMAP with a different (hyper-) parameter settings. -->

<!-- The visualization of the model generated for the selected layout in $9\text{-}D$ shows edges between the clusters (@fig-mnist-tri-proj (b2)). This provides evidence for the small separations between clusters. Furthermore, the data does not uniformally distributed along the clusters and therefore shows densed points (@fig-mnist-tri-proj (b2). Also, the clusters are non-linear shaped (@fig-mnist-tri-proj (b3).   -->

### 2024-09-07 paper.tex
Updated PBMC3k section

As shown in Figure~\ref{fig-NLDR-variety}, layout c, d, and e that show
small separations between clusters are universally optimal. But layout d
with little separations perform well at tiny binwidth and poorly as
binwidth increases. Also, layout e performs well at large binwidths. The
layout c is chosen as the best, since it used the UMAP with a different
(hyper-) parameter settings.

The visualization of the model generated for the selected layout in
\(9\text{-}D\) shows edges between the clusters
(Figure~\ref{fig-mnist-tri-proj} (b2)). This provides evidence for the
small separations between clusters. Furthermore, the data does not
uniformally distributed along the clusters and therefore shows densed
points (Figure~\ref{fig-mnist-tri-proj} (b2). Also, the clusters are
non-linear shaped (Figure~\ref{fig-mnist-tri-proj} (b3).

>>>>

As a result, it is necessary to find an appropriate layout for the
dataset. MSE for different binwidths (\(a_1\)) using tSNE, UMAP, PHATE,
PaCMAP, and TriMAP with various (hyper-)parameter settings were computed
(Figure~\ref{fig-pbmc-mse}). Layouts c, d, and e, which show small
separations between clusters, are universally optimal. However, layout d
performs well with smaller binwidths and poorly with larger binwidths.
On the other hand, layout e performs well with larger binwidths. Layout
c was selected for further analysis due to its better (hyper-)parameter
selection for the same method of the published plot.

The visualization of the selected layout in the \(9\text{-}D\) shows
edges between the clusters (Figure~\ref{fig-mnist-tri-proj} (b2)). This
supports the presence of small separations between clusters.
Additionally, the data points are not uniformly distributed across the
clusters, resulting in dense areas (Figure~\ref{fig-mnist-tri-proj}
(b2)). Furthermore, the clusters shows non-linear shapes
(Figure~\ref{fig-mnist-tri-proj} (b3)).
